---
title: "MDSC-206-Project"
author: "krishnakanth-20233"
date: "3/31/2021"
output: word_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries
```{r}
library(shinydashboard)
library(shiny)
library(heplots)
library(plyr)
library(e1071)
library(tidyverse)
library(caret)
library(gmodels)
library(gridExtra)
library(factoextra)
library(corrplot)
library(nortest)
library(MASS)
library(lattice)
library(broom)
library(cluster)
library(randomForest)
```

# Data setup
```{r}
data_excel <- read.csv("Spotify_data.csv")
set.seed(12)
data <- data_excel[sample(nrow(data_excel), 7000),]

dim(data)
head(data)
```

# Data cleaning

## checking for duplicate rows
```{r}
data[duplicated(data$names), ]
```

## checking for Missing values
```{r}
sum(is.na(data))
```

## Adding new variable using existing variables
```{r}
data <- data %>% mutate(duration_min = duration_ms/60000)
```

## Removing unnecessary columns
```{r}
summary(data)
data <- data[,-c(4,7,15)]
dim(data)
head(data)
```
 

## checking datatypes
```{r}
glimpse(data)
```
we can observe from above output the variables mode, explicit and key are integers but actually they are factors. So, convert them to factors.

## conversion to factors
```{r}
data$mode <- as.factor(data$mode)
data$key <- as.factor(data$key)
data$explicit <- as.factor(data$explicit)
glimpse(data)
```

# Data splitting
```{r}
set.seed(8)
dim(data)
index = sample(nrow(data), 5000)
train = data[index, ]
test = data[-index, ]

dim(train)
dim(test)
```

# Function for trends
```{r}
trends_plot <- function(a){
trend_change <- train %>% filter(year>2010) %>% group_by(year) %>% summarize_at(vars(all_of(a)), funs(Average = mean))
plot<- ggplot(data = trend_change, aes(x = year, y = Average)) +
  geom_line(color = "dodgerblue3", size = 1) +
  scale_x_continuous(breaks=seq(2011, 2020, 1))+
  scale_y_continuous(name=paste("",a,sep=""))
return(plot)
}
```

# linear model
```{r}
## Normalizing linear Model
energyn <- sqrt(train$energy)

## linear model after normalization
Nlinear_model <- lm(energyn ~ acousticness+loudness+valence ,data=train)
summary(Nlinear_model)

## plots
par(mfrow=c(2,2),mar=c(4,4,2,1))
plot(Nlinear_model)
```

## Predictions with Nlinear_model on train data
```{r}
trainPred <- predict(Nlinear_model,data = train,interval = "confidence",level = .99)
trainPred <- trainPred^2

train_df <- data.frame(cbind(actuals=train$energy, predicteds=trainPred))
head(train_df)
```

## Predictions with Nlinear_model on test data
```{r}
testPred <- predict(Nlinear_model,newdata = test,interval = "confidence",level = .99)
testPred <- testPred^2

test_df <- data.frame(cbind(actuals=test$energy, predicteds=testPred))
head(test_df)
```

# logistic model
```{r}
logistic_model = glm(explicit ~ acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min, data = train, family = binomial(link = "logit"))
summary(logistic_model)
```

## Predictions on test data

```{r}
test_pred <- predict(logistic_model, newdata = test, type = "response")
test_prediction <- ifelse(test_pred > 0.1, 1, 0)
test_log_mat <- table(predicted = test_prediction, actual = test$explicit)
confusionMatrix(test_log_mat, positive = "1")
```

## Naive Bayes model
```{r}
Naive_bayes <- naiveBayes(explicit ~ acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min, data = train)
Naive_bayes
```

## Predictions on test data
```{r}
test_pred <- predict(Naive_bayes, newdata = test)
test_Nbayes_mat <- table(predicted= test_pred, actual =test$explicit)
confusionMatrix(test_Nbayes_mat)
```

## QDA model
```{r}
qda_model <- qda(explicit ~acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min,train)
qda_model
```
## Predictions on test data
```{r}
test_pred <-  predict(qda_model, newdata = test)
test_qda_mat <- table(predicted = test_pred$class,actual = test$explicit)
confusionMatrix(test_qda_mat)
```

# Multinomial model
```{r}
require(nnet)
multinom_model <- multinom(explicit ~ acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min, data = train)
summary(multinom_model)
```
## Predictions on test data
```{r}
test_Pred <- predict(multinom_model, newdata = test, "class")
test_multi_mat <- table(predicted = test_Pred,actual = test$explicit)
confusionMatrix(test_multi_mat)
```

# kNN
```{r echo=FALSE}
train_ctrl <-trainControl(method = "cv", number = 8)
set.seed(10)
kNN <- train(explicit~ acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min,data = train, method ="knn",trControl = train_ctrl, preProcess = c("center", "scale"),tuneLength = 8)
kNN
```
kNN

## plots
```{r}
plot(kNN)
```

## Predictions on test data
```{r}
test_pred_kNN <- predict(kNN,newdata=test)
test_kNN_mat <- table(predicted = test_pred_kNN,actual = test$explicit)
confusionMatrix(test_kNN_mat, reference = test$explicit)
```

# Random forest
```{r}
forest <- randomForest(explicit~ acousticness+danceability+energy+loudness+popularity+speechiness+valence+year+duration_min,data= train,ntree = 600, mtry = 4, importance = TRUE)
forest
```

## Predicting on test set
```{r}
predTest <- predict(forest, test, type = "class")
test_forest_mat <- table(predTest, test$explicit)
confusionMatrix(test_forest_mat)
```

## Principal component analysis
```{r}
pc <- prcomp(train[,-c(2,5,7,10,11)],scale. = T)
summary(pc)
```

## scree plot
```{r}
plot(eigen_values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph')
lines(eigen_values)
```

# shiny code
```{r}

ui <- dashboardPage(
  skin = "blue",
  dashboardHeader(title = "Spotify Data Analysis"),
  dashboardSidebar(
    width = 300,
    sidebarMenu(
    menuItem("Introduction",tabName = "intro",icon = icon("spotify")),
    menuItem("Exploratory data analysis",
             menuSubItem("Data",tabName = "data",icon = icon("table")),
             menuSubItem("Correlation plot",tabName = "correlation_plot",icon = icon("check-circle")),
             menuSubItem("Trends",tabName = "Trends",icon = icon("chart-line")),
             menuSubItem("Histograms",tabName = "Histograms",icon = icon("chart-bar")),
             menuSubItem("Frequency plot",tabName = "Frequency_plot",icon = icon("creative-commons-sampling")),
             menuSubItem("Bar plot",tabName = "Bar_plot",icon = icon("bar-chart-o")),
             menuSubItem("Count plot",tabName = "count_plot",icon = icon("sort-amount-up-alt")),
             menuSubItem("Cross Tables",tabName = "cross_tables",icon = icon("table")),
             menuSubItem("Scatter plot",tabName = "scatter_plot",icon = icon("braille")),
             menuSubItem("Box plot",tabName = "box_plot",icon = icon("square")),icon = icon("book")),
    menuItem("PCA",
               menuSubItem("summary",tabName = "pca",icon = icon("list-alt")),
               menuSubItem("scree plot",tabName = "splot",icon = icon("chart-area")),icon = icon("envelope")),
     menuItem("K-means Clustering",
               menuSubItem("Clustering",tabName = "km",icon = icon("layer-group")),
               menuSubItem("Elbow curve",tabName = "eplot",icon = icon("chart-area")),icon = icon("envelope")),
     menuItem("Linear Model",
               menuSubItem("summary",tabName = "lm",icon = icon("list-alt")),
               menuSubItem("Plots",tabName = "plot",icon = icon("chart-area")),
               menuSubItem("predictions",tabName = "lmpred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("Logistic Model",
               menuSubItem("summary",tabName = "log",icon = icon("list-alt")),
               menuSubItem("predictions",tabName = "logpred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("Naive bayes Model",
               menuSubItem("summary",tabName = "nb",icon = icon("list-alt")),
               menuSubItem("predictions",tabName = "nbpred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("QDA",
               menuSubItem("summary",tabName = "qda",icon = icon("list-alt")),
               menuSubItem("predictions",tabName = "qdapred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("MLR",
               menuSubItem("summary",tabName = "mm",icon = icon("list-alt")),
               menuSubItem("predictions",tabName = "mmpred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("KNN",
               menuSubItem("summary",tabName = "knn",icon = icon("list-alt")),
               menuSubItem("Plot",tabName = "knnplot",icon = icon("chart-area")),
               menuSubItem("predictions",tabName = "knnpred",icon = icon("book-reader")),icon = icon("envelope")),
     menuItem("Random Forest",
               menuSubItem("summary",tabName = "rf",icon = icon("list-alt")),
               menuSubItem("predictions",tabName = "rfpred",icon = icon("book-reader")),icon = icon("envelope"))
    )
  ),
  dashboardBody(
    tabItems(
      tabItem("intro",
              plotOutput("img"),
              fluidRow(
              box(p("I was assigned the task of exploring any dataset that piqued my interest. I went for the Spotify dataset. Spotify is the most popular music streaming service on the planet. Users of the service need only register to gain access to one of the world's largest music libraries, as well as podcasts and other audio material. While exploring it in innumerable ways viz Exploratory Data Analysis, Linear models, Logistic models, etc., the inferences I accumulated were mind-boggling."),title = "Introduction", status = "success",width=8))
      ),
      tabItem("data",
              fluidPage(
                dataTableOutput("spotify_data")
              )
      ),
      tabItem("correlation_plot",
              box(selectInput("method", "Method:",
                c( "number",
                   "circle",
                   "square",
                   "pie")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
               box(width = 10,title = "Correlation plot", status = "primary",solidHeader = TRUE, plotOutput("corplot")),
      ),
      tabItem("Trends",
              box(selectInput("var", "Variable",
                c("acousticness",
                  "danceability",
                   "energy",
                   "instrumentalness",
                  "liveness",
                   "speechiness",
                   "popularity",
                   "loudness",
                   "valence",
                    "duration_min",
                   "tempo")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
              box(width = 10,title = "Trends over decade", status = "primary",solidHeader = TRUE, plotOutput("trends"))
      ),
      tabItem("Histograms",
              box(selectInput("var1", "Variable",
                c("acousticness",
                  "danceability",
                  "energy",
                  "instrumentalness",
                  "liveness",
                  "speechiness",
                  "popularity",
                  "loudness",
                  "valence",
                  "duration_min",
                  "year",
                  "tempo")),
                 sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
              box(width = 10,title = "Histogram", status = "primary",solidHeader = TRUE, plotOutput("hist"))
        ),
       tabItem("Frequency_plot",
               box(selectInput("var2", "Variable",
                c("acousticness",
                  "danceability",
                  "energy",
                  "instrumentalness",
                  "liveness",
                  "speechiness",
                  "popularity",
                  "loudness",
                  "valence",
                  "duration_min",
                  "year",
                  "tempo")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
                box(width = 10,title = "Frequency plot", status = "primary",solidHeader = TRUE, plotOutput("freq"))
       ),
      tabItem("Bar_plot",
               box(selectInput("var3", "Variable",
                c("mode","explicit","key")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
                box(width = 10,title = "Bar plot", status = "primary",solidHeader = TRUE, plotOutput("bar"))
       ),
      tabItem("count_plot",
               box(selectInput("var4", "Variable",
                c("mode","explicit","key")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
                box(width = 10,title = "Count plot", status = "primary",solidHeader = TRUE, plotOutput("count"))
       ),
      tabItem("cross_tables",
               fluidPage(box(selectInput("var10", "Variable",c("mode","explicit","key")),
                         selectInput("var11", "Variable",c("explicit","mode","key")),
                             title = "Input", status = "primary",solidHeader = TRUE,width = 3),
                box(width = 10,title = "Cross Table", status = "primary",solidHeader = TRUE, verbatimTextOutput("cross")))
       ),
       tabItem("scatter_plot",
               box(selectInput("var5", "X:",
                c("acousticness",
                  "danceability",
                  "energy",
                  "instrumentalness",
                  "liveness",
                  "speechiness",
                  "popularity",
                  "loudness",
                  "valence",
                  "duration_min",
                  "year",
                  "tempo")),
                selectInput("var6", "Y:",
                c("acousticness",
                  "danceability",
                  "energy",
                  "instrumentalness",
                  "liveness",
                  "speechiness",
                  "popularity",
                  "loudness",
                  "valence",
                  "duration_min",
                  "year",
                  "tempo")),title = "Input", status = "primary",solidHeader = TRUE,width = 3),
                box(width = 10,title = "Scatter plot", status = "primary",solidHeader = TRUE, plotOutput("scatter"))
       ),
      tabItem("box_plot",
              box(selectInput("var7", "Variable",
                c("acousticness",
                  "danceability",
                  "energy",
                  "instrumentalness",
                  "liveness",
                  "speechiness",
                  "popularity",
                  "loudness",
                  "valence",
                  "duration_min",
                  "year",
                  "tempo")),
                selectInput("opt","Option",c("No","Yes")),
                title = "Input", status = "primary",solidHeader = TRUE,
                 width = 3),
              p("*** Set option as yes to get Boxplots in comparision with Explicit "),
              box(width = 10,title = "Boxplot", status = "primary",solidHeader = TRUE, plotOutput("box"))
        ),
      tabItem("lm",h1("Linear model"),verbatimTextOutput("lm")),
      tabItem("plot",h1("Model plots"),plotOutput("plot")),
      tabItem("lmpred",h1("Model predictions on test data"),verbatimTextOutput("lmpred")),
      tabItem("log",h1("Logistic model"),verbatimTextOutput("logis")),
      tabItem("logpred",h1("Model predictions on test data"),verbatimTextOutput("logpred")),
      tabItem("nb",h1("Naive Bayes model"),verbatimTextOutput("nb")),
      tabItem("nbpred",h1("Model predictions on test data"),verbatimTextOutput("nbpred")),
      tabItem("qda",h1("Quadratic discrimination anaysis"),verbatimTextOutput("qda")),
      tabItem("qdapred",h1("Model predictions on test data"),verbatimTextOutput("qdapred")),
      tabItem("mm",h1("Multinominal logistic Regression"),verbatimTextOutput("mm")),
      tabItem("mmpred",h1("Model predictions on test data"),verbatimTextOutput("mmpred")),
      tabItem("knn",h1("K-Nearest Neibhours"),verbatimTextOutput("knn")),
      tabItem("knnplot",h1("Accuracy plots"),plotOutput("knnplot")),
      tabItem("knnpred",h1("Model predictions on test data"),verbatimTextOutput("knnpred")),
      tabItem("rf",h1("Random Forest"),verbatimTextOutput("rf")),
      tabItem("rfpred",h1("Model predictions on test data"),verbatimTextOutput("rfpred")),
      tabItem("pca",h1("Principal component analysis"),verbatimTextOutput("pca")),
      tabItem("splot",h1("Scree plot"),plotOutput("splot")),
      tabItem("km",h1("Cluster plot"),
              box(sliderInput(inputId = "k",
                  label = "k- value:",
                  min = 1,
                  max = 10,
                  value = 3)),
              plotOutput("km")),
      tabItem("eplot",h1("Elbow Curve"),plotOutput("eplot"))
    )#tabitems
  )
)

server <- function(input, output)
{
  output$img <- renderImage({filename <- normalizePath(file.path('./logo.png'))
    list(src = filename,width="780")
  }, deleteFile = FALSE)
  output$spotify_data <- renderDataTable(train)
  output$corplot <- renderPlot({corrplot(cor(train[,-c(2,5,7,10,11)]),method = input$method)})
  output$trends <- renderPlot({trends_plot(input$var)})
  output$hist <- renderPlot({
    ggplot(train,aes(train[,input$var1]))+geom_histogram(bins=input$bins,aes(y=..density..,fill=..count..)) +
       scale_fill_gradient("Count", low="aquamarine2", high="coral2")+
       geom_density()+xlab(input$var1)})
  output$freq <- renderPlot({
    ggplot(train, mapping = aes(train[,input$var2], colour = explicit )) +
  geom_freqpoly(binwidth = 0.1)+xlab(input$var2)})
  output$bar <- renderPlot({ ggplot(train, aes(train[,input$var3])) + geom_bar(aes(fill=train[,input$var3]))+xlab(input$var3)+ labs(fill = input$var3)})
  output$count <- renderPlot({ggplot(train) +geom_count(aes(train[,input$var4],key),color= "indianred2")+xlab(input$var4)})
  output$cross <- renderPrint({CrossTable(train[,input$var10],train[,input$var11])})
  output$scatter <- renderPlot({ggplot(train,aes(train[,input$var5],train[,input$var6]))+geom_point(col="lawngreen")+xlab(input$var5)+ylab(input$var6)})
  output$box <- renderPlot({
    if(input$opt=="No")
       ggplot(train,aes(train[,input$var7]))+geom_boxplot(color="slategrey",fill="turquoise")+xlab(input$var7)
    else
       ggplot(train,aes(train[,input$var7],train[,5]))+geom_boxplot(color="slategrey",fill=c("turquoise","lightcoral"))+xlab(input$var7)+ylab("Explicit")})
      
  output$lm <- renderPrint({summary(Nlinear_model)})
  output$plot <- renderPlot({par(mfrow=c(2,2),mar=c(4,4,2,1))
    plot(Nlinear_model)})
  output$lmpred <- renderPrint({head(test_df,15)})
  output$logis <- renderPrint({summary(logistic_model)})
  output$logpred <- renderPrint({confusionMatrix(test_log_mat, positive = "1")})
  output$nb <- renderPrint({Naive_bayes})
  output$nbpred <- renderPrint({confusionMatrix(test_Nbayes_mat)})
  output$qda <- renderPrint({qda_model})
  output$qdapred <- renderPrint({confusionMatrix(test_qda_mat)})
  output$mm <- renderPrint({summary(multinom_model)})
  output$mmpred <- renderPrint({confusionMatrix(test_multi_mat)})
  output$knn <- renderPrint({kNN})
  output$knnplot <- renderPlot({plot(kNN)})
  output$knnpred <- renderPrint({confusionMatrix(test_kNN_mat)})
  output$rf <- renderPrint({forest})
  output$rfpred <- renderPrint({confusionMatrix(test_forest_mat)})
  output$pca <- renderPrint({summary(pc)})
  output$splot <- renderPlot({plot(eigen_values, xlab = 'Eigenvalue Number', ylab = 'Eigenvalue Size', main = 'Scree Graph')
lines(eigen_values)})
  output$km <- renderPlot({clustering <- scale(data[,-c(2,5,7,10,11)])
    k <- kmeans(clustering, centers = input$k,nstart = 30)
    fviz_cluster(k, geom = "point",  data = clustering) + ggtitle("k=2")})
  output$eplot <- renderPlot({clustering <- scale(data[,-c(2,5,7,10,11)])
    fviz_nbclust(clustering, kmeans, method = "wss")})
}


shinyApp(ui = ui, server = server)
```

